---
title: "RLohiya Assignment12"
author: "Ritesh Lohiya"
date: "November 14, 2018"
output: html_document
---

#                             ASSIGNMENT 12

#           IS 605 FUNDAMENTALS OF COMPUTATIONAL MATHEMATICS

The attached who.csv dataset contains real-world data from 2008. The variables included follow.

Country: name of the country

LifeExp: average life expectancy for the country in years

InfantSurvival: proportion of those surviving to one year or more

Under5Survival: proportion of those surviving to five years or more

TBFree: proportion of the population without TB.

PropMD: proportion of the population who are MDs

PropRN: proportion of the population who are RNs

PersExp: mean personal expenditures on healthcare in US dollars at average exchange rate

GovtExp: mean government expenditures per capita on healthcare, US dollars at average exchange rate

TotExp: sum of personal and government expenditures.

```{r}
#who data
who <- read.csv('https://raw.githubusercontent.com/Riteshlohiya/Data605_HW12/master/who.csv')
summary(who)
```

1. Provide a scatterplot of LifeExp~TotExp, and run simple linear regression. Do not transform the variables. Provide and interpret the F statistics, R^2, standard error,and p-values only. Discuss whether the assumptions of simple linear regression met.

```{r}
#scatterplot
lm <- lm(LifeExp ~ TotExp, data = who)
plot(who$TotExp, who$LifeExp, xlab = 'Total Expenditure', ylab = 'Average Life Expectancy', main='Average Life Expectancy vs Expenditure')
abline(lm)
summary(lm)
```

```{r}
#plots
hist(resid(lm), main = "Histogram of Residuals", xlab = "Residuals")
plot(fitted(lm), resid(lm))
qqnorm(lm$residuals)
qqline(lm$residuals)
```

By seeing the R^2 value, we can say that the explained variability is very less @25.77%. The p-value suggests a statistically significant correlation between total expenditures and life expectancy, since p<<0.05. Looking at residuals plots it is clear that there is no constant variability and that residuals are not normally distributed. This model is not a good model.

2. Raise life expectancy to the 4.6 power (i.e., LifeExp^4.6). Raise total expenditures to the 0.06 power (nearly a log transform, TotExp^.06). Plot LifeExp^4.6 as a function of TotExp^.06, and r re-run the simple regression model using the transformed variables. Provide and interpret the F statistics, R^2, standard error, and p-values. Which model is "better?"

```{r}
#model2
LifeExp46 <- who$LifeExp**4.6
TotExp06 <- who$TotExp**0.06
lm1 <- lm(LifeExp46 ~ TotExp06, data = who)
summary(lm1)
plot(who$TotExp**0.06, who$LifeExp**4.6, xlab = 'Total Expenditure raised to 0.06 ', ylab = 'Average Life Expectancy raised to 4.6', main='Average Life Expectancy vs Expenditure')
abline(lm1)
```

```{r}
#plots
hist(resid(lm1), main = "Histogram of Residuals", xlab = "Residuals")
plot(fitted(lm1), resid(lm1))
qqnorm(lm1$residuals)
qqline(lm1$residuals)
```

The 2nd model is better than the 1st model. The R^2 is .729, that means the explained variability is 72.9% compared to 25% for the1st model. Looking at residuals plots, variability is fairly constant with a few outliers and distribution of residuals is nearly normal. So  2nd model is better than the 1st model.

3. Using the results from 3, forecast life expectancy when TotExp^.06 =1.5. Then forecast life expectancy when TotExp^.06=2.5.

The equation: 

y =  - 736527910 + 620060216*(x)

```{r}
LE <- function(x)
{y <- -736527910 + 620060216 *(x) 
y <- y^(1/4.6)
y}
LE(1.5)
LE(2.5)
```

4. Build the following multiple regression model and interpret the F Statistics, R^2, standard error, and p-values. How good is the model?

LifeExp = b0+b1 x PropMd + b2 x TotExp +b3 x PropMD x TotExp

```{r}
#model3
lm2 <- lm(LifeExp ~ PropMD + TotExp + PropMD*TotExp, data = who)
summary(lm2)
```

```{r}
#plots
hist(resid(lm2), xlab = "Residuals")
plot(fitted(lm2), resid(lm2))
```

The R^2 is just .3574 that means that the explained variability is just 35.74%. Residuals is right skewed and do not show constant variability. Dont think this is good model.

5. Forecast LifeExp when PropMD=.03 and TotExp = 14. Does this forecast seem realistic? Why or why not?

```{r}
LifeExp_new <- ( (6.277*10^1) + (1.497*10^3)*.03 + (7.233*10^(-5))*14 - ((6.026*10^(-3))*0.03*14) ) 
LifeExp_new
```

This seems like an outlier as life expectancy of 107 does not look correct.
